---
title: "Practical Machine Learning (Project)"
author: "TD"
date: "Monday, February 16, 2015"
output: pdf_document
---


##Introduction
In this project, we will use the data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to find patterns in their behavior. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: (http://groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset). 

##Synopsis
In this project, two methods were used to fit the training data set. Cross validation were applied. Based on the model evaluation, it reveals that model using the method of random forest is better than the model using the method of 'rpart' in terms of accuracy of prediction. Finally, the random forest model were applied to predict the new test data set (Test_final).


## Load data sets

```{r cache=TRUE}
#setwd("E:\\Coursera\\Practical Machine Learning\\project")
#UrlTrain <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
#UrlTest <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
#download.file (UrlTrain, destfile = "training.csv")
#download.file (UrlTest, destfile = "testing.csv")

setwd("E:\\Coursera\\Practical Machine Learning\\project")
setTrain<-read.table(file="training.csv", header = TRUE, sep = ",")
setTest<-read.table(file="testing.csv", header = TRUE, sep = ",")
```

##Preprocess data sets

####Keep variables that satisfying the following criteria:
1. variables that indicate measures from accelerometers on the belt, forearm, arm, and dumbell
2. no missing values
3. no near zero variables

```{r cache=TRUE, message=FALSE}
dim(setTrain);dim(setTest)

#select the variables that relates to 
#the data from accelerometers on the belt, forearm, arm, and dumbell
library(dplyr)
Train<-select(setTrain, contains("accel"),raw_timestamp_part_1,raw_timestamp_part_2, classe)
names(Train)
Test<-select(setTest, contains("accel"),raw_timestamp_part_1,raw_timestamp_part_2)
names(Test)

#check missing data and keep complete columns
VarMtest<-which(colSums(is.na(Test))>0)
Test_final<-Test[,-c(VarMtest)]
include<- names(Train) %in% c("classe", names(Test[,-c(VarMtest)])) 
Train_final <- Train[include]

#check near zero variables
library(caret)
nsv <- nearZeroVar(Train_final, saveMetrics=TRUE)
nsv 

#double check
dim(Train_final);dim(Test_final)


sapply(Train_final, function(x) sum(is.na(x) | x == ""))
sum(!complete.cases(Train_final))
sapply(Test_final, function(x) sum(is.na(x) | x == ""))
sum(!complete.cases(Test_final))
```

####Split the 'Train_final' set into 'training' set(60%) and 'testing' set(40%)

```{r cache=TRUE, message=FALSE}
set.seed(1968)
library(caret)
inTrain = createDataPartition(y=Train_final$classe, p=0.6, list=FALSE)
training = Train_final[inTrain,]
testing = Train_final[-inTrain,]
dim(training);dim(testing)

summary(training$classe)
```

##Fit Models

####Model1: Predicting with trees using 'rpart'
```{r cache=TRUE, message=FALSE}
library(rattle)
library(rpart.plot)

model<-train(classe~., method="rpart", data=training)
#print(model,digits=3)
#print(model$finalModel, digits=3)
```

####Model2: Predicting with random forest
```{r cache=TRUE, message=FALSE}
library(randomForest)
model2<- randomForest(classe~., data=training,proximity=TRUE)
#print(model2,digits=3)
```


##Model evaluation

####Model1 (method="rpart")
```{r cache=TRUE, message=FALSE}
predictions<-predict(model, newdata=testing)
confusionMatrix(predictions, testing$classe)
```

####Model2 (method="randomForest")
```{r cache=TRUE, message=FALSE}
predictions2<-predict(model2, newdata=testing, type="class")
confusionMatrix(predictions2, testing$classe)
```

##Prediction on the 'Test_final' set

```{r cache=TRUE, message=FALSE}
predict_new<-predict(model2, newdata=Test_final, type="class")
predict_new
```

##Appendix: Tree plot of the final model that using the method of 'rpart'
```{r fig.width=8, fig.height=6, cache=TRUE}
fancyRpartPlot(model$finalModel)
```

